# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: Madi Subaiti

========================================================

# Initialization

Import libraries

```{r}
library(data.table)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(tidyverse)
library(magrittr)

```



Load the critics dataset.

```{r}
folder_path <- "/Users/madisubaiti/Desktop/BIG DATA/TERM 2/Recommendation Engines/IE-Lab1/Data"
critics <- fread(file.path(folder_path, "critics.csv"));
critics;
View(critics)
```




# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Sophia and Nuria


```{r fig.width=7, fig.height=6, eval=T, echo=F}
#1---

#Making DF and Transposing the dataset  
critics_x <- as.data.frame(critics[, -1])
critics_t <- t(critics_x)
# Assigning names to the column since it is not numerica has been deleted
colnames(critics_t) <- t(critics[,1])
rownames(critics_x) <- t(critics[,1])

# Correlating Nuria and Sophia using cor function.
PCC1 <- as.data.frame(na.omit(critics_t[,c("Nuria", "Sophia")]))
print(cor(PCC1[,1],PCC1[,2]))




#b)--- [don't run] -----

# Transpose the dataset
crit2<-t(critics[,2:21]) 

# Since they were names not numeric values need to give them names
colnames(crit2)<-t(critics[,1]) # Assign names to columns

#correlate
cor(crit2,use="complete.obs") 

cor(crit2,use="pairwise.complete.obs") # use pairwise deletion

crit2<-as.data.frame(crit2)
with(crit2, cor(Sophia, Nuria, use = "complete.obs", method = "pearson"))



#-------------------
#**Expected correlation between Sophia and Nuria**: 
0.3651484

```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Sophia and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?
  
```{r fig.width=7, fig.height=6}
  
#2----
PCC2 <- as.data.frame(na.omit(critics_t[,c("Maria", "Nerea")]))
print(cor(PCC1[,1],PCC1[,2]))

PCC3 <- as.data.frame(na.omit(critics_t[,c("Chris", "Jim")]))
print(cor(PCC1[,1],PCC1[,2]))

#Scatter plot for Sophia and Nuria
plot1 = PCC1 %>%
  ggplot(aes(x=Nuria, y=Sophia)) +
  geom_point(col="blue") +
  labs(x="Nuria Ratings", y="Sophia Ratings", title="Correlation of Sophia vs. Nuria")+
  geom_text_repel(aes(label=rownames(PCC1))) + 
  coord_fixed(1, xlim = c(1, 5), ylim = c(1,5))
  print(plot1);
  

# Scatter plot of Maria and Nerea
plot2 = PCC2 %>%
  ggplot(aes(x=Maria, y=Nerea)) +
  geom_point(col="red") +
  labs(x="Maria Ratings", y="Nerea Ratings", title="Correlation of Maria vs. Nerea")+
  geom_text_repel(aes(label=rownames(PCC2))) + 
  coord_fixed(1, xlim = c(1, 5), ylim = c(1,5))
  print(plot2);

# Scatter plot of Chris and Jim

plot3 = PCC3 %>%
  ggplot(aes(x=Chris, y=Jim)) +
  geom_point(col="purple") +
  labs(x="Chirs Ratings", y="Jim Ratings", title="Correlation of Chris vs. Jim")+
  geom_text_repel(aes(label=rownames(PCC3))) + 
  coord_fixed(1, xlim = c(1, 5), ylim = c(1,5))
  print(plot3);




```  
  

# 3) Top Recommendations 
Return the top-5 recommended movies for Sophia Use the weighted average of every other critic's rankings.



```{r fig.width=7, fig.height=6}
#3----

# creating distict columns for movies, ratings and weighted averages from each user rating 
critics1 <- critics %>%
  gather("movie", "rating", -User) %>%
  group_by(User) %>%
  mutate(mean = mean(rating, na.rm = TRUE)) %>%
  ungroup()

# Listing movies Sophia hasn´t seen or rated by filtering out her ratings
sophia_movies_not_seen <- critics1 %>%
  filter(User == "Sophia" & is.na(rating)) %>%
  extract2("movie")

# Finding Sophia´s movie ratings 
sophia_ratings <- critics1 %>%
  filter(User == "Sophia") %>%
  extract2("rating")

# Correlate between Sophia´s ratings and the rest by creating new variables eliminating the nulls
sophia_corr <- critics1 %>%
  group_by(User) %>%
  mutate(
    pearson = cor(rating, sophia_ratings, use = "complete.obs"),
    pearson_abs = abs(pearson)
  ) %>%
  ungroup()

# The movies recommended to Sophia filtering to movies not rated by her
sophia_recom <- sophia_corr %>%
  filter(movie %in% sophia_movies_not_seen) %>%
  group_by(movie) %>%
  summarise(
    score = sum((rating - mean) * pearson, na.rm = TRUE) / (sum(pearson_abs) - 1)) %>%
  mutate(score = score + mean(sophia_ratings, na.rm = TRUE)) %>%
  arrange(desc(score)) 
  
  head(sophia_recom,5)


#------
#**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'
```

# 4) Top similar critics
Return the top 5 critics that are most similar to Sophia. 



```{r fig.width=7, fig.height=6}
#4 ----

# variable for critic
critic <-"Sophia"

#getting the position of sophia
position <- grep(critic, rownames(critics_x))
no_rate <- c(is.na(critics_x[position,]))

#  an empty vector to iterate the correlations of sophia with other users

correlations<- c()
for (i in 1:ncol(critics_t))
  {
  PCC1 <-na.omit(as.data.frame(cbind(critics_t[,i],critics_t[,critic])))
  correlations[i] <- cor(PCC1[,1],PCC1[,2])
  ;
}

corr_df <- data.frame(correlation = correlations[-position],
                              row.names = rownames(critics_x[-position,]))


head(corr_df[order(-corr_df$correlation), , drop = FALSE],5)





#-----------
**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'
```